Deep Reinforcement Learning for Conservation Decisions

### 2021-01-22 manuscript/manuscript.Rmd
Adding algorithm table

Many reinforcement learning algorithms center around estimating value functions, which attempt to quantify the goodness of being in a state or taking an action from a state. Given that the objective of RL is to maximize the cumulative expected reward, "goodness" refers to the likelihood that a state or state-action pair will result in a high reward trajectory. There are 3 value functions that are commonly used across RL algorithms: the state value function, $V^\pi(s_t)$, the state-action value function, $Q^\pi(s_t, a_t)$, and the advantage function, $A^\pi(s_t, a_t)$. The state-value function is defined as the expected return from a given state,
$$
  V^\pi(s_t) = \mathbb{E}_{\tau \sim p_\pi(\tau \mid s_t)}\Bigg[\sum_{t' = t}^H \gamma^{t' - t} r(s_t, a_t)\Bigg]
$$
Meanwhile, the state-action-value or quality function, finds the expected return given a state and action,
$$
  Q^\pi(s_t, a_t) = \mathbb{E}_{\tau \sim p_\pi(\tau \mid s_t, a_t)}\Bigg[\sum_{t' = t}^H \gamma^{t' - t} r(s_t, a_t)\Bigg]
$$
And lastly, the advantage function characterizes a relative state-action value in that it quantifies how valuable a state-action pair is over a baseline. The advantage function can then be succintly defined as
$$
  A^\pi(s_t, a_t) = Q^\pi(s_t, a_t) - V^\pi(s_t)
$$
Learning $Q^\pi$, commonly called the Q-value function, is more powerful than learning the state-value function as Q-values easily prescribe what action to take when in a certain state, $a_t^* = \underset{a_t}{\text{argmax}} \, Q^\pi (s_t, a_t)$. Conversely, if we know the state-value function, we would have to perform a maximization over an expected value to find the optimal action, $a_t^* = \underset{a_t}{\text{argmax}} \, \mathbb{E}_{s_{t+1} \sim \,T(s_{t+1} \mid s_t, a_t)}[V^\pi(s_{t+1})]$, which is less tractable. Yet, learning the advantage function is often preferred in practice as advantage estimators favorably have lower variance than state-value and Q-value estimators [cite schulman gae].

@@

The field of reinforcement learning has proliferated over the past 10 years on account of contributions from deep learning. In every reinforcement learning algorithm, the agent is attempting to learn some function, whether it is a policy function and/or a value function. Neural networks, which have the property of being universal function approximators [cite], are thus a powerful tool in reinforcement learning since neural networks can be used to approximate policy and/or value functions which we do not know the form of. Additionally, the ability of neural networks to autonomously devise low-dimensional representatoins of data is very beneficial when state observations are high dimension, for instance in the case of image-based observations. The typical process for using neural networks in RL is that the agent is attempting to optimize one or more neural networks using the backpropagation algorithm -- an exception is made here for evolutionary algorithms which did not employ gradient-based methods. Deep RL has excelled in solving sequential decision-making problems that have non-tabular state-action spaces, problems that are typically intractable to classic approaches.

In practice, the general process to attaining a good performing agent is to tune model hyperparameters then train an agent using the best hyperparameters. Hyperparameters refer to the parameters in a RL algorithm that are not directly learned during the training process. For example, a weight in the neural network is a model parameter, but the number of hidden layers in the neural network is a hyperparameter. Since the optimal hyperparameters cannot be found analytically, tuning should be done to find a good-performing set of hyperparameters. The hyperparameter tuning process consists of sampling then evaluating a set of hyperparameters, where tuning trials are often ranked by mean cumulative reward. It is important to note that hyperparameter tuning can often be a very time intensive process. Since one round of training can be time intensive, hyperparameter tuning, which consists of usually numerous rounds of training, can be very time intensive.

(Big table listing all the algorithms)

>>>>

Many reinforcement learning algorithms center around estimating value functions, which attempt to quantify the goodness of being in a state or taking an action from a state. Given that the objective of RL is to maximize the cumulative expected reward, "goodness" refers to the likelihood that a state or state-action pair will result in a high reward trajectory. There are 3 value functions that are commonly used across RL algorithms: the state value function, $V^\pi(s_t)$, the state-action value function, $Q^\pi(s_t, a_t)$, and the advantage function, $A^\pi(s_t, a_t)$. The state-value function is defined as the expected return from a given state,
$$
  V^\pi(s_t) = \mathbb{E}_{\tau \sim p_\pi(\tau \mid s_t)}\Bigg[\sum_{t' = t}^H \gamma^{t' - t} r(s_t, a_t)\Bigg]
$$

Meanwhile, the state-action-value or quality function, finds the expected return given a state and action,
$$
  Q^\pi(s_t, a_t) = \mathbb{E}_{\tau \sim p_\pi(\tau \mid s_t, a_t)}\Bigg[\sum_{t' = t}^H \gamma^{t' - t} r(s_t, a_t)\Bigg]
$$

And lastly, the advantage function characterizes a relative state-action value in that it quantifies how valuable a state-action pair is over a baseline. The advantage function can then be succintly defined as
$$
  A^\pi(s_t, a_t) = Q^\pi(s_t, a_t) - V^\pi(s_t)
$$
Learning $Q^\pi$, commonly called the Q-value function, is more powerful than learning the state-value function as Q-values easily prescribe what action to take when in a certain state, $a_t^* = \underset{a_t}{\text{argmax}} \, Q^\pi (s_t, a_t)$. Conversely, if we know the state-value function, we would have to perform a maximization over an expected value to find the optimal action, $a_t^* = \underset{a_t}{\text{argmax}} \, \mathbb{E}_{s_{t+1} \sim \,T(s_{t+1} \mid s_t, a_t)}[V^\pi(s_{t+1})]$, which is less tractable. Yet, learning the advantage function is often preferred in practice as advantage estimators favorably have lower variance than state-value and Q-value estimators [cite schulman gae].

@@

The field of reinforcement learning has proliferated over the past 10 years on account of contributions from deep learning. In every reinforcement learning algorithm, the agent is attempting to learn some function, whether it is a policy function and/or a value function. Neural networks, which have the property of being universal function approximators [cite], are thus a powerful tool in reinforcement learning since neural networks can be used to approximate policy and/or value functions which we do not know the form of. Additionally, the ability of neural networks to autonomously devise low-dimensional representatoins of data is very beneficial when state observations are high dimension, for instance in the case of image-based observations. The typical process for using neural networks in RL is that the agent is attempting to optimize a policy and/or value network using the backpropagation algorithm -- an exception is made here for evolutionary algorithms which did not employ gradient-based methods. Deep RL has excelled in solving sequential decision-making problems that have non-tabular state-action spaces, problems that are typically intractable with classical methods.

In practice, the general process for attaining a good performing agent is to tune model hyperparameters then train an agent using the best hyperparameters. Hyperparameters refer to the parameters in a RL algorithm that are not directly learned during the training process. For example, a weight in the neural network is a model parameter, but the number of hidden layers in the neural network is a hyperparameter. Since the optimal hyperparameters cannot be found analytically, tuning should be done to find a good-performing set of hyperparameters. The hyperparameter tuning process consists of sampling then evaluating a set of hyperparameters, where tuning trials are often ranked by mean cumulative reward. It is important to note that hyperparameter tuning can often be a very time intensive process. Since one round of training can be time intensive, hyperparameter tuning, which consists of usually numerous rounds of training, can be very time intensive.

```{=latex}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
Algorithm & Description & Model & Policy & Action Space & State Space \\
\hline
MBPO & Model-based Policy Optimization & Model-based & On-policy& &\\
\hline
MCTS & Monte Carlo Tree Search & Model-based & Either & & \\
\hline
A2C  & Advantage Actor Critic & Model-free & On-policy& &\\
\hline
A3C  & Asynchronous A2C & Model-free & On-policy & &\\
\hline
TRPO  & Trust Region Policy Optimization & Model-free & On-policy & &\\
\hline
PPO  & Proximal Policy Optimization & Model-free & On-policy & &\\
\hline
DQN & Deep Q Networks & Model-free & Off-policy& &\\
\hline
DDPG & Deep Deterministic Policy Gradient & Model-free & Off-policy & &\\
\hline
TD3 & Twin Delayed DDPG & Model-free & Off-policy & &\\
\hline
SAC & Soft Actor Critic & Model-free & Off-policy & &\\
\hline
IMPALA & Importance Weighted Actor Learner & Model-free & Off-policy & &\\
\hline
APE-X & Distributed Prioritized Experience Replay & Model-free & Off-policy & & \\
\hline
\end{tabular}
\end{center}
```
